---
title: "Final Project"
author: "Ethan McCollister"
date: "10/17/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE)
library(tidyverse)
library(janitor)
library(knitr)

hitters_2018 <- read_csv("hitters_2018.csv")
hitters_2017 <- read_csv("hitters_2017.csv")
hitters_2016 <- read_csv("hitters_2016.csv")
hitters_2015 <- read_csv("hitters_2015.csv")
hitters_2014 <- read_csv("hitters_2014.csv")
hitters_2013 <- read_csv("hitters_2013.csv")
hitters_2012 <- read_csv("hitters_2012.csv")

pitchers_2018 <- read_csv("pitchers_2018.csv") %>%
  clean_names()
pitchers_2017 <- read_csv("pitchers_2017.csv") %>%
  clean_names()
pitchers_2016 <- read_csv("pitchers_2016.csv") %>%
  clean_names()
pitchers_2015 <- read_csv("pitchers_2015.csv") %>%
  clean_names()
pitchers_2014 <- read_csv("pitchers_2014.csv") %>%
  clean_names()
pitchers_2013 <- read_csv("pitchers_2013.csv") %>%
  clean_names()
pitchers_2012 <- read_csv("pitchers_2012.csv", col_types = cols(`vFO (pfx)` = col_double(),
                                                                `FO-X (pfx)` = col_double(),
                                                                `FO-Z (pfx)` = col_double())) %>%
  clean_names()
```

```{r adding_year_column}
pitchers_2018 <- pitchers_2018 %>% 
  mutate(year = "2018")
pitchers_2017 <- pitchers_2017 %>% 
  mutate(year = "2017")
pitchers_2016 <- pitchers_2016 %>% 
  mutate(year = "2016")
pitchers_2015 <- pitchers_2015 %>% 
  mutate(year = "2015")
pitchers_2014 <- pitchers_2014 %>% 
  mutate(year = "2014")
pitchers_2013 <- pitchers_2013 %>% 
  mutate(year = "2013")
pitchers_2012 <- pitchers_2012 %>% 
  mutate(year = "2012")
```

## Preci

I currently have two data frames uploaded to my repo - hitters.csv and pitchers.csv. I am realizing as I write this that these are poorly named, as it is actually just data from 2018, and I intend to bring in data for every year that it's available (I think since 2010). Eventually I will have one large dataframe for hitters data and one for pitchers. 

hitters.csv contains plate discipline data (swing%, contact%, swinging strike%, etc.), and pitchers.csv contains Pitch f/x data (velocity, pitch usage, movement, etc.) My goal for the project is to employ clustering techniques to try to create groupings of hitters and groupings of pitchers, then compare them to see if certain categories of one have particular success against certain categories of the other.

Below we see the names of the variables in the hitters.csv dataframe, as well as the number of rows.
```{r preci}
names(hitters)

nrow(hitters)

```

## Creating Master Dataframes
```{r creating_masters}
hitters_all <- hitters_2018 %>%
  full_join(hitters_2017) %>%
  full_join(hitters_2016) %>%
  full_join(hitters_2015) %>%
  full_join(hitters_2014) %>%
  full_join(hitters_2013) %>%
  full_join(hitters_2012)

pitchers_all <- pitchers_2018 %>%
  full_join(pitchers_2017) %>%
  full_join(pitchers_2016) %>%
  full_join(pitchers_2015) %>%
  full_join(pitchers_2014) %>%
  full_join(pitchers_2013) %>%
  full_join(pitchers_2012)
```

```{r}
write_csv(pitchers_all, "pitchers_all.csv")

```



